%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------


\documentclass{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{comment}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsfonts}

\hyphenation{op-tical net-works semi-conduc-tor}

\input{cfg-commands}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}


\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Making the heap symbolic in symbolic execution }
%\subtitle{Subtitle Text, if any}

\authorinfo{Anonymous}
          {}
    %       {Email1}
%\authorinfo{Name2\and Name3}
 %         {Affiliation2/3}
  %       {Email2/3}

\maketitle

\begin{abstract}
One of the fundamental challenges of using symbolic execution
to analyze software has been the treatment of dynamically
allocated data.
%, i.e., programs with complex data structures with
%unbounded domains.
State-of-the-art symbolic execution techniques 
have addressed this challenge by constructing the heap \emph{lazily},
materializing objects on the concrete heap ``as needed'' and
using non-deterministic choice points to explore each feasible
concrete heap configuration. Because analysis of the materialized
heap locations relies on concrete program semantics, the lazy
initialization approach exacerbates the state space explosion
problem that limits the scalability of symbolic execution.
In this work we present a novel approach for lazy symbolic execution
of heap manipulating software which utilizes 
a fully symbolic heap constructed on-the-fly during symbolic execution.
Our approach is 1) \emph{scalable} -- it does not
create the additional points of non-determinism
introduced by existing lazy initialization techniques and it explores
each execution path only once for any given set of isomorphic heaps,
2) \emph{precise} -- at any given point during symbolic execution, 
the symbolic heap represents the exact set of feasible
concrete heap structures for the program under analysis, and
3) \emph{expressive} -- the symbolic heap can represent recursive data structures 
and heaps resulting from loops and recursive control structures in the code. 
We report on a case-study of an implementation of our technique in the
Symbolic PathFinder tool to illustrate its scalability, precision
and expressiveness. We also discuss how test case generation -- a common
 use for symbolic execution results -- can benefit from symbolic
 execution which uses a fully symbolic heap.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2

\input{intro}
\input{motivation}
\input{pseudo-code}
\input{proofs}
%\input{stateDescriptions}
\input{related}
%\input{state}
%\input{execution}
%\input{examples}
%\input{bytecode}
%\section{Conclusion}

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{../bib/paper}


\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices

