\section{Rules for Javalite}\label{app:javalite}

\input{javalite-common}

The surface syntax and the machine syntax for the Javalite syntactic
machine are in the main paper. \figref{fig:javalite-common} defines
the rewrite relations for the portion of the Javalite language
semantics that are common to both generalized symbolic execution and
the symbolic heap algorithm. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking or error conditions. The
machine simply halts if no rewrite is enabled.

%% Includes its section definition
\input{lazy}	

\input{precise-fig}

\input{summarize-precise}

\section{Symbolic Heap Algorithm}


%\begin{figure*}[t]
%\begin{center}
%\setlength{\tabcolsep}{60pt}
%\hspace*{-35pt}
%\begin{tabular}[c]{cc}
%\scalebox{1.0}{\usebox{\boxPFAFW}} & 
%%\scalebox{0.91}{\input{faYHeap.pdf_t}} &
%\scalebox{0.91}{\input{fwXHeap.pdf_t}} \\ \\
%(a) & (b)
%\end{tabular}
%\end{center}
%\caption{Field read and write relations with an example heap. (a) Field-access, $\rsym^\mathit{A}$, and field-write, $\rsym^\mathit{W}$, rewrite rules for the $\rsym$ relation. (b) The final heap after $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\ \cfgt{:=}\ \lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp\rp$.}
%\label{fig:fHeap}
%\end{figure*}

%\begin{figure*}
%\begin{tabular}[c]{l}
%\scalebox{1.0}{\usebox{\boxPI}} \\
%\end{tabular}
%\caption{The summary machine, $s ::= \lp\cfgnt{L}\ \cfgnt{R}\ \cfgnt{r}\ \cfgnt{f}\ \cfgnt{C}\rp$, with $s\rsum^*s^\prime =  s \rsum \cdots \rsum s^\prime \rsum s^\prime$.}
%\label{fig:symInit}
%\end{figure*}

The rewrite rules for field access or field write when a null reference is possible leads to an error state.
$$
\begin{array}{c}
	\inferrule[Field Access (NULL)]{
      \exists \lp \phi\ l\rp \in \cfgnt{L}\lp \cfgnt{r}\rp\ \lp l = l_{\mathit{null}} \wedge \mathbb{S}\lp \phi \wedge \phi_g\rp \rp
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgnt{r}\ \lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp \rp  \rsym^\mathit{A^\prime}
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgt{error}\ \cfgt{end} \rp
	} \\
\\
	\inferrule[Field Write (NULL)]{
      \cfgnt{r}_x = \eta\lp \cfgnt{x}\rp \\
      \exists \lp \phi\ l\rp \in \cfgnt{L}\lp \cfgnt{r}_x\rp\ \lp l \neq l_{\mathit{null}} \wedge \mathbb{S}\lp \phi \wedge \phi_g\rp \rp
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgnt{r}\ \lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgt{*}\ \rightarrow\ \cfgnt{k}\rp \rp  \rsym^\mathit{F^\prime}
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgt{error}\ \cfgt{end}\rp
	}	
\end{array}
$$ The rules reflect a branch in the control flow for the symbolic
heap algorithm: one branch being the feasible null outcome (shown
here), and the other branch being the potential non-null outcome
(shown in the main paper). In the non-null case, the path constraint
is updated to restrict out all possible null instances.

%\input{definitions}


